---
title: "DATA 550: Final Project Report"
author: "Dan Flanagan"
date: "`r Sys.Date()`"
output: html_document
params:
  production: TRUE
---

```{r, here-i-am, include = FALSE}
here::i_am(
  "DF_report.Rmd"
)
```

```{r, set-chunk-opts, echo = FALSE, eval = params$production}
library(knitr)
opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE
)

```

```{r, load-pkgs}
library(dplyr)
library(table1)
library(ggplot2)

```

```{r, load-data}
here::i_am("code/DF_DATA550_FinalProj2.Rmd")

push_data <- readRDS(here::here("data/push_tbeval.rds"))

```

## Introduction

For the final project, I am planning to use data from the PUSH trial - a randomized clinical trial of children with HIV in Kenya. The dataset includes observations (rows) for `r nrow(push_data)` children with HIV who were hospitalized for concurrent illness and evaluated for TB disease. The dataset includes variables for clinical, radiographic, diagnostic tests, and biomarker data for each participant.

The objectives for this analysis are to: 

- Summarize descriptive characteristics of the PUSH cohort
- Assess data completeness and distribution of key diagnostic and outcome variables
- (Potentially) Conduct multivariable logistic regression of clinical, radiographic, and laboratory data to predict TB disease


## Descriptive Characteristics

```{r, eda_push, echo=FALSE, include = FALSE}
n_distinct(push_data$study_id) #168 participants

summary(push_data$sex_en) #Categorical Var (1=Male, 2=Female)
  table(push_data$sex_en)
summary(push_data$enroll_ageyrs) #Continuous Var (years)
summary(push_data$cgeduc_en) #Categorical Var (NEED VALUES)
  table(push_data$cgeduc_en)
summary(push_data$resnum_en) #Categorical Var (# of people in home)
  table(push_data$resnum_en)
summary(push_data$roomnum_en) #Categorical Var (# of rooms in home)
  table(push_data$roomnum_en)
summary(push_data$primarymother) #Dichotomous Var (primary caregiver is mother; 0-No, 1=Yes)
  table(push_data$primarymother)
  
summary(push_data$immbcgbirth_en) #Dichotomous Var (0=No, 1=Yes)
  table(push_data$immbcgbirth_en)
summary(push_data$priorptb) #Dichotomous Var (0=No, 1=Yes)
  table(push_data$priorptb) 
summary(push_data$priorhosp_en) #Dichotomous Var (0=No, 1=Yes)
  table(push_data$priorhosp_en) 
summary(push_data$priorhospnum_en) #Continuous Var (# of hospitalizations)
  table(push_data$priorhospnum_en)

summary(push_data$tbcont_en) #Dichotomous Var (0=No, 1=Yes)
  table(push_data$tbcont_en)  
  
#tb outcomes
summary(push_data$tbclass_new) #Categorical (1=Confirmed, 2=Unconfirmed, 3=Unlikely)
  table(push_data$tbclass_new)


```


```{r, prep-table-one, include=FALSE}

##assign factor labels to categorical/dichotomous variables of interest
push_data$sex_en <- factor(push_data$sex_en, levels=c(1,2), labels=c("Male", "Female"))
push_data$tbclass_new <- factor(push_data$tbclass_new, levels=c(1,2,3), labels=c("Confirmed", "Unconfirmed", "Unlikely"))

##label variable headings for table1
label(push_data$sex_en) <- "Participant Sex"
label(push_data$enroll_ageyrs) <- "Age at Enrollment"
units(push_data$enroll_ageyrs) <- "years"
label(push_data$tbclass_new) <- "TB Disease"
strata <- c(list(Overall=push_data), split(push_data, push_data$tbclass_new))
labels <- list(
            variables=list(sex_en=render.varlabel(push_data$sex_en),
                           enroll_ageyrs=render.varlabel(push_data$enroll_ageyrs)),
            groups=list("","TB Disease"))

my.render.cont <- function(x) {
    with(stats.default(x), c("",
        "Mean (SD)"=sprintf("%0.1f (%0.1f)", MEAN, SD)))
}

my.render.cat <- function(x) {
    c("", sapply(stats.default(x), function(y) with(y,
        sprintf("%d (%0.0f%%)", FREQ, PCT))))
}

caption <- "Table 1. Basic demographics of PUSH trial participants (n=181) by TB Disease Status"
footnote <- c("Continuous variables reported as Mean (SD).", 
              "Categorical variables reported as n (%). Values may not sum to total (N) due to missing data.")


```

```{r, table-one, include = TRUE}
table1(strata, labels, groupspan = c(1,3), render.missing=NULL,
       render.continuous=my.render.cont,
       render.categorical=my.render.cat,
       caption=caption,
       footnote=footnote)
```



## Age Distribution of PUSH Trial Participants

```{r, include = TRUE}
## Plot age distribution of PUSH study participants
ggplot(push_data, aes(x = enroll_ageyrs)) + 
  geom_histogram(binwidth = 1, fill = "steelblue", color = "white") + 
  labs( title = "Age Distribution of Study Participants",
    x = "Age (years)",
    y = "Count"
  ) +
  theme_minimal()


```


```{r, include = FALSE}
## Plot age distribution of PUSH participants, stratified by age
ggplot(push_data, aes(x = enroll_ageyrs, fill=sex_en)) + 
  geom_histogram(binwidth = 1, color = "white", alpha = 0.7, position = "stack") + 
  labs( title = "Age Distribution of Study Participants",
    x = "Age (years)",
    y = "Count",
    fill = "Sex"
  ) +
  theme_minimal()


```




