---
title: "DATA 550: Final Project Report"
author: "Dan Flanagan"
date: "`r Sys.Date()`"
output: html_document
params:
  production: TRUE
---

```{r, here-i-am, include = FALSE}
here::i_am(
  "DF_report.Rmd"
)
```

```{r, set-chunk-opts, echo = FALSE, eval = params$production}
library(knitr)
opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE
)

```

```{r, load-pkgs}
library(knitr)
library(pandoc)

```

```{r, load-data}
push_data <- readRDS(here::here("data/push_sim_data.rds"))

# load pre-generated table1 object
table1_obj <- readRDS(here::here("output/table1.rds"))

# load pre-generated figure1 object
figure1_obj <- readRDS(here::here("output/figure1.rds"))

```

## Introduction

For the final project, I am using a simulated data set from the PUSH trial - a randomized clinical trial of children with HIV in Kenya. The dataset includes observations (rows) for `r nrow(push_data)` children with HIV who were hospitalized for concurrent illness and evaluated for TB disease. The dataset includes variables for clinical, radiographic, diagnostic tests, and biomarker data for each participant.

The objectives for this analysis are to: 

- Summarize descriptive characteristics of the PUSH cohort
- Assess data completeness and distribution of key diagnostic and outcome variables


## Descriptive Characteristics

```{r, table-one, include = TRUE}
table1_obj

```



## Age Distribution of PUSH Trial Participants

```{r, figure-one, include = TRUE}
figure1_obj

```




